---
title: "Optimal Predictor Selection for High-dimensional Nonparametric Forecasting"
author: Nuwani Palihawadana
titlegraphic: bg-13.png
titlecolor: white
toc: true
format:
  presentation-beamer:
    pdf-engine: xelatex
    template-partials:
        - before-title.tex
execute:
  cache: true
  echo: true
  warning: false
  message: false
---

```{r}
#| label: load-packages
#| echo: false
library(knitr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tsibble)
library(lubridate)
library(cowplot)
library(patchwork)
library(kableExtra)
library(feasts)
```


# Sparse Multiple Index (SMI) Model

## Heat Exposure Related Daily Mortality
```{r}
#| label: heat-summer-plot
#| echo: false
#| message: false

dataSummer <- readRDS(here::here("data/Heat_Corrected.rds"))
dataSummer |>
  as_tsibble(index = Date) |>
  mutate(Day = row_number()) |>
  update_tsibble(index = Day, regular = TRUE) |>
  autoplot(Death_lag_000, colour = "#D55E00") +
  scale_x_continuous(
    breaks = seq(1, 2300, by = 92),
    labels = unique(dataSummer$Year)
  ) +
  labs(x = "Date", y = "Number of Deaths", title = "Daily Deaths in Summer - Montreal, Canada") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 12, size = 7))
```


## Heat Exposure Related Daily Mortality
```{r}
#| label: temp-heat-plot
#| echo: false
#| message: false

p1 <- dataSummer |>
  ggplot(aes(x = Tmax_lag_000, y = Death_lag_000)) +
  geom_point(colour = "#CC79A7", alpha = 0.3) +
  xlim(c(0, 35)) +
  labs(
    x = "Maximum Temperature (Degrees Celsius)",
    y = "Number of Deaths",
    title = "Deaths (Summer) against Max Temperature"
  ) +
  theme_bw() +
  geom_smooth(colour = "black", se = FALSE, size = 0.4)

p2 <- dataSummer |>
  ggplot(aes(x = Tmin_lag_000, y = Death_lag_000)) +
  geom_point(colour = "#CC79A7", , alpha = 0.3) +
  xlim(c(0, 35)) +
  labs(
    x = "Minimum Temperature (Degrees Celsius)",
    y = "Number of Deaths",
    title = "Deaths (Summer) against Min Temperature"
  ) +
  theme_bw() +
  geom_smooth(colour = "black", se = FALSE, size = 0.4)

plot_grid(p1, p2, nrow=1)
```


## Background

\begin{itemize}
  \item \color{violet} \textbf{\textsl{Nonlinear "Transfer Function" model}}
\end{itemize}
\begin{block}{}
  \centerline{$y_{t} = f(\mathbfit{x}_{t}, \mathbfit{x}_{t-1}, \dots,\mathbfit{x}_{t-k}, y_{1},\dots,y_{t-l}) + \varepsilon_{t}$}
\end{block}

$y_{t}$ -- variable to forecast \newline
$\mathbfit{x}_{t}$ -- a vector of predictors \newline
$\varepsilon_{t}$ -- random error

\pause
\begin{itemize}
  \item Impossible to estimate $f$ for large $k$ -- \textbf{\textit{curse of dimensionality}}
  \pause
  \item Reasonable to impose additivity constraints \\[0.2cm]
\end{itemize}
\begin{block}{}
  \centerline{$f(\mathbfit{x}_{t}, \mathbfit{x}_{t-1}, \dots,\mathbfit{x}_{t-k}) = \sum_{a = 0}^{k} {f_{a}(\mathbfit{x}_{t-a})}$ \pause \color{violet} \alert{$\mathbfit{\leftarrow}$ Nonparametric Additive Model}}
\end{block}


## Background

::: {.callout-warning}
## \color{violet} \textbf{Issues:}

\begin{enumerate}
  \item Challenging to estimate in a high-dimensional setting \newline
  \item Subjectivity in predictor selection, and predictor grouping to model interactions
\end{enumerate}
:::


## Background

::: {.callout-warning}
## \color{violet} \textbf{Issues:}

\begin{enumerate}
  \item Challenging to estimate in a high-dimensional setting \newline
  \item Subjectivity in predictor selection, and predictor grouping to model interactions
\end{enumerate}
:::

::: {.callout-note}
## \color{violet} \textbf{Index Models:}

\begin{itemize}
    \item \textbf{Mitigate difficulty of estimating a nonparametric component for each predictor}
    \item \textbf{Improve flexibility}
\end{itemize}
\begin{textblock}{6}(8.5, 7)
\fontsize{12}{12}\sf
\begin{block}{}
\color{violet}
$$
  y_{i} = g \left ( \mathbfit{\alpha}^{T} \mathbfit{x}_{i} \right ) + \varepsilon_{i}
$$
\end{block}
\end{textblock}
:::


## Sparse Multiple Index (SMI) Model

\begin{block}{Semi-parametric model}
$$y_{i} = \beta_{0} + \sum_{j = 1}^{p}g_{j}(\mathbfit{\alpha}_{j}^{T}\mathbfit{x}_{ij}) + \sum_{k = 1}^{d}f_{k}(w_{ik}) + \mathbfit{\theta}^{T}\mathbfit{u}_{i} + \varepsilon_{i}, \quad i = 1, \dots, n,$$
\end{block}

\begin{itemize}
  \item \small \color{black} $y_{i}$ -- \color{violet} univariate response
  \item \small \color{black} $\mathbfit{x}_{ij} \in \mathbb{R}^{\ell_{j}}$, $j = 1, \dots, p$ -- \color{violet} $p$ subsets of predictors entering indices
  \item \small \color{black} $\mathbfit{\alpha}_{j}$ -- \color{violet} $\ell_{j}$-dimensional vectors of index coefficients
  \item \small \color{black} $g_{j}, f_{k}$ -- \color{violet} smooth nonlinear functions
  \item \small \color{black} Additional predictors :
    \begin{itemize}
      \item \small $w_{ik}$  --   \color{violet} nonlinear 
      \item \small \color{black} $\mathbfit{u}_{i}$  --   \color{violet} linear
    \end{itemize}
\end{itemize}

\only<2->{\begin{textblock}{5.5}(9, 6.3)
\fontsize{11}{12}\sf
\begin{alertblock}{}
\small Allow elements equal to zero in $\mathbfit{\alpha}_{j}$ -- "Sparse"
\end{alertblock}
\end{textblock}}

\only<3->{\begin{textblock}{5.5}(9, 6)
\fontsize{11}{12}\sf
\begin{alertblock}{}
\scriptsize Both "p" and the predictor grouping among indices are unknown.
\end{alertblock}
\end{textblock} 

\begin{textblock}{5.5}(9, 7)
\fontsize{11}{12}\sf
\begin{alertblock}{}
\scriptsize Overlapping of predictors among indices is not allowed.
\end{alertblock}
\end{textblock}}


## Optimisation Problem

\small Let \color{violet}\textbf{$q$} \color{black}be the \color{violet}\textit{total number of predictors} \color{black}entering indices.

\begin{block}{}
\begin{align*}
  & \min_{\beta_{0}, p, \mathbfit{\alpha}, \mathbfit{g}, \mathbfit{f}, \mathbfit{\theta}} \quad \sum_{i = 1}^{n}\Bigg [ y_{i} - \beta_{0} - \sum_{j = 1}^{p}g_{j}(\mathbfit{\alpha}_{j}^{T}\mathbfit{x}_{i}) - \sum_{k = 1}^{d}f_{k}(w_{ik}) - \mathbfit{\theta}^{T}\mathbfit{u}_{i}\Bigg]^{2} \\
  & \hspace*{5cm} + \color{blue}\lambda_{0}\sum_{j = 1}^{p}\sum_{m = 1}^{q}\mathbb{1}(\alpha_{jm} \neq 0) \color{black}+ \color{magenta}\lambda_{2}\sum_{j = 1}^{p}\|\mathbfit{\alpha}_{j}\|_2^2 \nonumber \\
  & \hspace*{1.5cm} \text{s.t.}\quad \sum_{j=1}^p \mathbb{1}(\alpha_{jm} \neq 0) \in \{0,1\} \quad \forall m \nonumber
\end{align*}
\end{block}
\begin{itemize}
  \item \color{blue}$\lambda_{0} > 0$ -- controls the number of selected predictors
  \item \color{magenta}$\lambda_{2} \ge 0$ -- controls the strength of the additional shrinkage
\end{itemize}


## MIQP Formulation

\begin{block}{}
$$
\begin{aligned}
  \min_{\beta_{0}, p, \mathbfit{\alpha}, \mathbfit{g}, \mathbfit{f}, \mathbfit{\theta}, \mathbfit{z}} \quad & \sum_{i = 1}^{n}\Bigg [ y_{i} - \beta_{0} - \sum_{j = 1}^{p}{g_{j}(\mathbfit{\alpha}_{j}^{T}\mathbfit{x}_{i})} - \sum_{k = 1}^{d} {f_{k}(w_{ik})} - \mathbfit{\theta}^{T}\mathbfit{u}_{i}\Bigg ]^{2} \\
  & \hspace*{4cm} + \color{blue}\lambda_{0}\sum_{j = 1}^{p}\sum_{m = 1}^{q}z_{jm} \color{black}+ \color{magenta}\lambda_{2}\sum_{j = 1}^{p}\sum_{m = 1}^{q} \alpha_{jm}^{2} \\
  \text{s.t.} \quad & |\alpha_{jm}| \le Mz_{jm} \quad \forall j, \forall m, \\
  & \sum_{j = 1}^{p}z_{jm} \le 1 \quad \forall m, \\
  & z_{jm} \in \{0, 1\} \color{red} \quad \mathbfit{\leftarrow} \quad z_{jm} = \mathbb{1}(\alpha_{jm} \neq 0)
\end{aligned}
$$
\end{block}
\begin{itemize}
  \item $M < \infty$: \small{If $\mathbfit{\alpha^{*}}$ is an optimal solution, then $\max \big ( \{|\alpha_{jm}^{*} |\}_{j\in[p],m\in[q]} \big) \le M$}
\end{itemize}


## Estimation Algorithm

\begin{block}{}
\fontsize{9}{9}\sf
\color{blue} \textbf{Step 1 :} \color{black} Initialise index structure and index coefficients
\end{block}
\pause
\begin{block}{}
\fontsize{8}{8}\sf
  \begin{itemize}
    \item \color{violet} \textbf{PPR:} Projection Pursuit Regression Based Initialisation
    \item \color{violet} \textbf{Additive:} Nonparametric Additive Model Based Initialisation
    \item \color{violet} \textbf{Linear:} Linear Regression Based Initialisation
    \item \color{violet} \textbf{Multiple:} Pick One From Multiple Initialisations
  \end{itemize}
\end{block}


## Estimation Algorithm

\begin{block}{}
\fontsize{9}{9}\sf
\color{blue} \textbf{Step 1 :} \color{black} Initialise index structure and index coefficients
\end{block}
\begin{block}{}
\fontsize{9}{9}\sf
\color{violet} \textbf{Step 2 :} \color{black} Estimate nonlinear functions
\end{block}
\pause
\begin{block}{}
\fontsize{9}{9}\sf
\color{blue} \textbf{Step 3 :} \color{black} Update index coefficients
\end{block}
\pause
\begin{block}{}
\fontsize{9}{9}\sf
\color{violet} \textbf{Step 4 :} \color{black} Iterate steps 2 and 3 -- until:
\end{block}
\pause
\begin{block}{}
\fontsize{8}{8}\sf
\begin{itemize}
  \item \color{violet} convergence
  \item \color{violet} loss increases for 3 consecutive iterations \color{black} OR
  \item \color{violet} max iterations
\end{itemize}
\end{block}


## Estimation Algorithm

\begin{block}{}
\fontsize{9}{9}\sf
\color{blue} \textbf{Step 1 :} \color{black} Initialise index structure and index coefficients
\end{block}
\begin{block}{}
\fontsize{9}{9}\sf
\color{violet} \textbf{Step 2 :} \color{black} Estimate nonlinear functions
\end{block}
\begin{block}{}
\fontsize{9}{9}\sf
\color{blue} \textbf{Step 3 :} \color{black} Update index coefficients
\end{block}
\begin{block}{}
\fontsize{9}{9}\sf
\color{violet} \textbf{Step 4 :} \color{black} Iterate steps 2 and 3 until stopping criteria are reached
\end{block}
\begin{block}{}
\fontsize{9}{9}\sf
\color{blue} \textbf{Step 5 :} \color{black} Add a new index with dropped predictors, and repeat step 4
\end{block}
\pause 
\begin{block}{}
\fontsize{9}{9}\sf
\color{violet} \textbf{Step 6 :} \color{black} Increase p by 1 in each iteration of step 5 -- until:
\end{block}
\pause 
\begin{block}{}
\fontsize{8}{8}\sf
\begin{itemize}
  \item \color{violet} no.of indices reaches $q$
  \item \color{violet} loss increases after the increment model \color{black} OR 
  \item \color{violet} solution maintains same no.of indices as previous iteration, and abs(difference of index coefficients between two successive iterations) $<=$ tolerance
\end{itemize}
\end{block}


## Forecasting Heat Exposure Related Daily Mortality

\begin{textblock}{7}(0.7, 1)
\fontsize{11}{12}\sf
\begin{block}{Variables}
  \begin{itemize}
    \item \color{violet} \textbf{Response:} \color{black} \textbf{Daily deaths in Summer} -- 1990 to 2014 -- Montreal, Canada
    \item \color{violet} \textbf{Index Variables:} 
      \begin{itemize}
        \item \color{black} Death lags
        \item \color{black} Max temperature lags
        \item \color{black} Min temperature lags
        \item \color{black} Vapor pressure lags
      \end{itemize}
    \item \color{violet}\textbf{Nonlinear:} \color{black} DOS (day of the season), Year \newline
  \end{itemize}
\end{block}
\end{textblock}

\only<2->{\begin{textblock}{14.6}(0.7, 6.8)
\fontsize{11}{12}\sf
\begin{alertblock}{}
$$
  \textbf{Deaths} = \beta_{0} + \sum_{j = 1}^{p}{g_{j}(\mathbfit{X}\mathbfit{\alpha}_{j})} + f_{1}(\textbf{DOS}) + f_{2}(\textbf{Year})+ \mathbfit{\varepsilon},
$$
\end{alertblock}
\end{textblock}}


## Forecasting Heat Exposure Related Daily Mortality

\begin{textblock}{7}(0.7, 1)
\fontsize{11}{12}\sf
\begin{block}{Variables}
  \begin{itemize}
    \item \color{violet} \textbf{Response:} \color{black} \textbf{Daily deaths in Summer} -- 1990 to 2014 -- Montreal, Canada
    \item \color{violet} \textbf{Index Variables:} 
      \begin{itemize}
        \item \color{black} Death lags
        \item \color{black} Max temperature lags
        \item \color{black} Min temperature lags
        \item \color{black} Vapor pressure lags
      \end{itemize}
    \item \color{violet}\textbf{Nonlinear:} \color{black} DOS (day of the season), Year \newline
  \end{itemize}
\end{block}
\end{textblock}

\begin{textblock}{7}(8.3, 1)
\fontsize{11}{12}\sf
\begin{block}{Data Split}
  \begin{itemize}
  \item \color{violet}\textbf{Training Set:} \color{black}1990 to 2012 \newline
  \item \color{violet}\textbf{Validation Set:} \color{black}2013 \newline
  \item \color{violet}\textbf{Test Set:} \color{black}2014 \newline \newline \newline \newline \newline \newline
\end{itemize}
\end{block}
\end{textblock}

\begin{textblock}{14.6}(0.7, 6.8)
\fontsize{11}{12}\sf
\begin{alertblock}{}
$$
  \textbf{Deaths} = \beta_{0} + \sum_{j = 1}^{p}{g_{j}(\mathbfit{X}\mathbfit{\alpha}_{j})} + f_{1}(\textbf{DOS}) + f_{2}(\textbf{Year})+ \mathbfit{\varepsilon},
$$
\end{alertblock}
\end{textblock}


## Results

\fontsize{9}{12}\sf
```{r}
#| echo: false
results_heat <- readr::read_csv(here::here("results/heat_results_correct.csv"))
kable(results_heat,
    format = "latex",
    booktabs = TRUE,
    digits = 3,
    escape = FALSE,
    linesep = "",
    col.names = c("Model", "Predictors", "Indices", "MSE", "MAE", "MSE", "MAE")
  ) |>
  add_header_above(c("", "", "", "Test Set 1" = 2, "Test Set 2" = 2), align = "c") |>
  kable_styling(latex_options = c("repeat_header")) |>
  row_spec(0, align = "c") |>
  column_spec(4, bold = if_else(results_heat$MSE1 == min(results_heat$MSE1), TRUE, FALSE)) |>
  column_spec(5, bold = if_else(results_heat$MAE1 == min(results_heat$MAE1), TRUE, FALSE)) |>
  column_spec(6, bold = if_else(results_heat$MSE2 == min(results_heat$MSE2), TRUE, FALSE)) |>
  column_spec(7, bold = if_else(results_heat$MAE2 == min(results_heat$MAE2), TRUE, FALSE))
```

\begin{textblock}{14}(1, 6.2)
\fontsize{10}{12}\sf
  \alert{SMI Model (a, b) $\quad \mathbfit{\rightarrow \quad \lambda_{0} = a, \lambda_{2} = b}$}
\end{textblock}

\begin{textblock}{14}(1, 7.2)
\fontsize{10}{12}\sf
\begin{block}{}
\begin{itemize}
  \item \color{violet} \textbf{Test Set 1:} \color{black} Three months (June, July and August 2014)
  \item \color{violet} \textbf{Test Set 2:} \color{black} One month (June 2014)
\end{itemize}
\end{block}
\end{textblock}


## Summary

::: {.callout-note}
## \color{violet} \textbf{Key features:}

\begin{itemize}
    \item Automatic selection of number of indices and predictor grouping
    \item Automatic predictor selection
    \item A wide spectrum: from single index models to additive models
  \end{itemize}
:::

::: {.callout-warning}
## \color{violet} Limitations:

\begin{itemize}
    \item \small Initialisation: we encourage trial-and-error
    \item \small Computational cost: increases with number of predictors and indices
  \end{itemize}
:::

\begin{textblock}{13.8}(1.1, 7.6)
\fontsize{11}{12}\sf
\begin{alertblock}{}
\color{blue} \hspace*{1.5cm} \textbf{Paper :} \href{https://github.com/nuwani-palihawadana/smimodel_paper}{github.com/nuwani-palihawadana/smimodel\_paper}
\end{alertblock}
\end{textblock}


## R Package - smimodel

\placefig{11.7}{2}{width=4cm}{hex-large}

\begin{textblock}{10.2}(1, 0.8)
\fontsize{12}{14}\sf
\begin{block}{}
  \begin{itemize}
    \item Open source implementation of \textbf{SMI Modelling Algorithm}
    \begin{itemize}
      \item \color{violet} \textbf{model\_smimodel()} \newline
    \end{itemize}
    \item Penalty parameter tuning with greedy search
    \begin{itemize}
      \item \color{violet} \textbf{greedy\_smimodel()} \newline
    \end{itemize}
    \item Functions to fit benchmark models
    \begin{itemize}
      \item \color{violet} \textbf{model\_backward()}
      \item \color{violet} \textbf{model\_gaim()}
      \item \color{violet} \textbf{model\_ppr() etc.}
    \end{itemize}
  \end{itemize}
\end{block}
\end{textblock}

\begin{textblock}{10.2}(1, 7.7)
\fontsize{11}{12}\sf
\begin{alertblock}{}
 \color{blue} \hspace*{0.8cm} \href{https://github.com/nuwani-palihawadana/smimodel}{github.com/nuwani-palihawadana/smimodel}
\end{alertblock}
\end{textblock}


# Uncertainty Estimation

## Uncertainty Estimation

\begin{itemize}
  \item \textbf{\color{violet}"Uncertainty"} of a forecast \alert{$\mathbfit{\rightarrow}$ Prediction Interval (PI)}
  \pause
  \item Theoretical $100(1 - \alpha)\%$ prediction interval:
$$
  \hat{y}_{t+h|t} \pm z_{\alpha/2}*\hat{\sigma}_{h},
$$
where
  \begin{itemize}
    \item \small \color{black} $y$ -- \color{violet} time series $y_{1}, \dots, y_{T}$
    \item \small \color{black} $\hat{y}_{t+h|t}$ -- \color{violet} $h$ steps ahead point forecast for $y_{t+h}$
    \item \small \color{black} $z_{\alpha/2}$ -- \color{violet} $\alpha/2$ quantile of standard normal distribution
    \item \small \color{black} $\hat{\sigma}_{h}$ -- \color{violet} an estimate of std. deviation of $h$-step forecast distribution
  \end{itemize}
  \pause
  \item Nonparametric Additive Models:
  \begin{itemize}
    \item \small \color{blue} No distributional assumptions
    \item \small \color{blue} Serially correlated errors \alert{$\mathbfit{\rightarrow}$ Impossible to estimate theoretical PIs} 
  \end{itemize}
\end{itemize}


## Background

\begin{block}{Block Bootstrap}
\begin{itemize}
  \item Resampling from empirical distribution of historical model residuals \color{violet} $\mathbfit{\rightarrow}$ Bootstrapping \newline
  \pause
  \item \color{black} Randomly resample blocks from the historical model residuals, and join together \alert{$\mathbfit{\rightarrow}$ Block Bootstrapping} 
  \item Retains serial correlation in the data 
  \pause
  \item \color{violet} block length:
  \begin{itemize}
    \item Long enough to capture autocorrelation patterns
    \item Short enough to construct sufficient number of blocks
  \end{itemize}
\end{itemize}
\end{block}


## Background

\begin{block}{\color{yellow} Conformal Prediction (CP) -- Vovk et al. (2005)}
\begin{itemize}
  \item A distribution-free approach
  \item Relies only on the assumption of \textbf{exchangeability of data}
  \item Provides theoretical coverage guarantees
\end{itemize}
\end{block}

\pause

\begin{block}{Split Conformal Prediction (SCP)}
\begin{itemize}
  \item A holdout method for generating prediction intervals
  \begin{itemize}
  \item \textbf{Training set} -- forecasting model is trained
  \item \textbf{Calibration set} -- forecasting errors (\textit{nonconformity scores}) are calculated
  \item \textbf{Test set} -- prediction intervals are obtained
  \end{itemize}
\end{itemize}
\end{block}


## Background

\begin{itemize}
  \item \color{violet}CP methods for \textbf{non-exchangeable data}:
\end{itemize}
\pause
\begin{block}{Weighted Conformal Prediction (WCP) Methods}
\begin{itemize}
  \item \color{violet}Tibshirani et al. (2019):
  \begin{itemize}
    \item Depends on \textbf{"covariate shift"} assumption
    \item \textit{Nonconformity scores} are weighted using ratio of likelihoods of training and test covariate distributions
    \item Likelihood ratio is assumed to be known or accurately estimated
  \end{itemize}
  \pause
  \item \color{violet}Barber et al. (2023):
  \begin{itemize}
    \item Weighting \textit{quantiles} to avoid assumption of exchangeability
    \item Weights are "fixed" rather than being data dependent
  \end{itemize}
\end{itemize}
\end{block}


## Background

\begin{itemize}
  \item \color{violet}CP methods for \textbf{non-exchangeable data}:
\end{itemize}
\begin{block}{Adaptive Conformal Prediction (ACP) -- Gibbs \& Candès (2021)}
\begin{itemize}
  \item Update nominal $\mathbfit{\alpha}$ based on achieved coverage
  \begin{itemize}
    \item If achieved coverage is larger -- increase $\mathbfit{\alpha}$
    \item If achieved coverage is smaller -- decrease $\mathbfit{\alpha}$ \newline
  \end{itemize}
\end{itemize}
\end{block}


## Methodology

\begin{block}{}
  \begin{itemize}
    \item Prediction interval construction methods \newline
    \begin{itemize}
      \item Block Bootstrapping (BB) \newline
      \item Conformal Prediction (CP) methods: SCP, WSCP, ACP \newline
    \end{itemize}
  \end{itemize}
\end{block}

\begin{block}{}
  \begin{itemize}
    \item Applied using \textit{\color{violet}online learning framework} proposed by \textit{\color{violet}Wang and Hyndman (2024)} \newline
  \end{itemize}
\end{block}


## Forecasting Heat Exposure Related Daily Mortality

\begin{textblock}{7}(0.7, 1.5)
\fontsize{11}{12}\sf
\begin{block}{Data Recap}
  \begin{itemize}
    \item \color{violet} \textbf{Response:} \color{black} \textbf{Daily deaths in Summer} -- 1990 to 2014 -- Montreal, Canada
    \item \color{violet} \textbf{Index Variables:} 
      \begin{itemize}
        \item \color{black} Death lags
        \item \color{black} Max temperature lags
        \item \color{black} Min temperature lags
        \item \color{black} Vapor pressure lags
      \end{itemize}
    \item \color{violet}\textbf{Nonlinear:} \color{black} DOS (day of the season), Year \newline
  \end{itemize}
\end{block}
\end{textblock}

\only<2->{\begin{textblock}{7}(8.3, 1.5)
\fontsize{11}{12}\sf
\begin{block}{\color{yellow} New Data Split}
  \begin{itemize}
  \item \color{violet}\textbf{Training Set:} \color{black}1990 to 2007 \newline
  \item \color{violet}\textbf{Validation Set:} \color{black}2008 \newline
  \item \color{violet}\textbf{Test Set:} \color{black}2009 to 2014 \newline \newline \newline \newline \newline \newline
\end{itemize}
\end{block}
\end{textblock}}


## Results

\color{violet} \textbf{Block Bootstrap} $95\%$ Prediction Intervals (block length = 59):
```{r}
#| label: fig-int-bb
#| echo: false
#| message: false

intervals <- readRDS(here::here("results/intervals_combine_new.rds"))
cols_interval <- c(
  "true" = "grey40",
  "lower" = "#009E73",
  "upper" = "#D55E00"
)
intervals |>
  mutate(Horizon = if_else(Horizon == 1, "H = 1", if_else(Horizon == 46, "H = 46", "H = 92"))) |>
  filter(Method == "BB") |>
  ggplot(aes(x = Time, y = Value, group = Type, colour = Type)) +
  labs(
    y = "Number of Deaths"
    ) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  facet_grid(factor(Model, levels = c("SMI - PPR", "Backward", "GAIM", "PPR")) ~ Horizon) +
  scale_colour_manual(values = cols_interval) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 12),
        legend.box = "vertical",
        legend.title = element_blank(),
        axis.text.x = element_text(size = 7)) +
  xlab("Time Index (t)") +
  ylim(c(0, 150))
```


## Results

\color{violet} \textbf{Split Conformal Prediction} $95\%$ Prediction Intervals:
```{r}
#| label: fig-int-scp
#| echo: false
#| message: false

intervals |>
  mutate(Horizon = if_else(Horizon == 1, "H = 1", if_else(Horizon == 46, "H = 46", "H = 92"))) |>
  filter(Method == "SCP") |>
  ggplot(aes(x = Time, y = Value, group = Type, colour = Type)) +
  labs(
    y = "Number of Deaths"
    ) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  facet_grid(factor(Model, levels = c("SMI - PPR", "Backward", "GAIM", "PPR")) ~ Horizon) +
  scale_colour_manual(values = cols_interval) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 12),
        legend.box = "vertical",
        legend.title = element_blank(),
        axis.text.x = element_text(size = 7)) +
  xlab("Time Index (t)") +
  ylim(c(0, 150))
```


## Results

\color{violet} \textbf{Weighted Split Conformal Prediction} $95\%$ Prediction Intervals:
```{r}
#| label: fig-int-wscp
#| echo: false
#| message: false

intervals |>
  mutate(Horizon = if_else(Horizon == 1, "H = 1", if_else(Horizon == 46, "H = 46", "H = 92"))) |>
  filter(Method == "WSCP") |>
  ggplot(aes(x = Time, y = Value, group = Type, colour = Type)) +
  labs(
    y = "Number of Deaths"
    ) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  facet_grid(factor(Model, levels = c("SMI - PPR", "Backward", "GAIM", "PPR")) ~ Horizon) +
  scale_colour_manual(values = cols_interval) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 12),
        legend.box = "vertical",
        legend.title = element_blank(),
        axis.text.x = element_text(size = 7)) +
  xlab("Time Index (t)") +
  ylim(c(0, 150))
```


## Results

\color{violet} \textbf{Adaptive Conformal Prediction} $95\%$ Prediction Intervals:
```{r}
#| label: fig-int-acp
#| echo: false
#| message: false

intervals |>
  mutate(Horizon = if_else(Horizon == 1, "H = 1", if_else(Horizon == 46, "H = 46", "H = 92"))) |>
  filter(Method == "ACP") |>
  ggplot(aes(x = Time, y = Value, group = Type, colour = Type)) +
  labs(
    y = "Number of Deaths"
    ) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  facet_grid(factor(Model, levels = c("SMI - PPR", "Backward", "GAIM", "PPR")) ~ Horizon) +
  scale_colour_manual(values = cols_interval) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 12),
        legend.box = "vertical",
        legend.title = element_blank(),
        axis.text.x = element_text(size = 7)) +
  xlab("Time Index (t)") +
  ylim(c(0, 150))
```


## Results

\color{violet} \textbf{Mean Coverage:}
```{r}
#| label: fig-cov
#| echo: false
#| message: false

avgCov_data_smimodel_ppr <- readRDS(here::here("results/SMIModel_PPR/plotData_combined_avgCov_heat_smimodel_ppr.rds"))
avgCov_data_backward <- readRDS(here::here("results/Backward/plotData_combined_avgCov_heat_backward.rds"))
avgCov_data_gaim <- readRDS(here::here("results/GAIM/plotData_combined_avgCov_heat_gaim.rds"))
avgCov_data_ppr <- readRDS(here::here("results/PPR/plotData_combined_avgCov_heat_ppr.rds"))
methods <- c("BB", "SCP", "WSCP", "ACP")
cols <- c(
  "BB" = "black",
  "SCP" = "#0072B2",
  "WSCP" = "#009E73",
  "ACP" = "#D55E00"
)

p1 <- avgCov_data_smimodel_ppr |>
  mutate(method = factor(method, levels = methods)) |>
  ggplot(aes(x = horizon, y = coverage, group = method, colour = method)) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  scale_colour_manual(values = cols) +
  geom_hline(yintercept = 0.95, linetype = "dashed", colour = "gray40") +
  xlab("Forecast horizon") +
  ylab("Mean coverage") +
  ylim(c(0.80, 1.00)) +
  ggtitle("SMI Model - PPR") +
  theme_bw() +
  theme(axis.title.x = element_blank()) +
  scale_x_discrete(breaks = seq(1, 92, by = 5))

p2 <- avgCov_data_backward |>
  mutate(method = factor(method, levels = methods)) |>
  ggplot(aes(x = horizon, y = coverage, group = method, colour = method)) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  scale_colour_manual(values = cols) +
  geom_hline(yintercept = 0.95, linetype = "dashed", colour = "gray40") +
  xlab("Forecast horizon") +
  ylab("Mean coverage") +
  ylim(c(0.80, 1.00)) +
  ggtitle("Backward") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  scale_x_discrete(breaks = seq(1, 92, by = 5))

p3 <- avgCov_data_gaim |>
  mutate(method = factor(method, levels = methods)) |>
  ggplot(aes(x = horizon, y = coverage, group = method, colour = method)) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  scale_colour_manual(values = cols) +
  geom_hline(yintercept = 0.95, linetype = "dashed", colour = "gray40") +
  xlab("Forecast horizon") +
  ylab("Mean coverage") +
  ylim(c(0.80, 1.00)) +
  ggtitle("GAIM") +
  theme_bw() +
  scale_x_discrete(breaks = seq(1, 92, by = 5))

p4 <- avgCov_data_ppr |>
  mutate(method = factor(method, levels = methods)) |>
  ggplot(aes(x = horizon, y = coverage, group = method, colour = method)) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  scale_colour_manual(values = cols) +
  geom_hline(yintercept = 0.95, linetype = "dashed", colour = "gray40") +
  xlab("Forecast horizon") +
  ylab("Mean coverage") +
  ylim(c(0.80, 1.00)) +
  ggtitle("PPR") +
  theme_bw() +
  theme(axis.title.y = element_blank()) +
  scale_x_discrete(breaks = seq(1, 92, by = 5))

patchwork <- (p1 + p2) / (p3 + p4) & theme(legend.position = "bottom", 
                                           #legend.box = "vertical",
                                           legend.title = element_blank(),
                                           axis.text.x = element_text(size = 7),
                                           axis.title = element_text(size = 8),
                                           title = element_text(size = 8))
patchwork +
  plot_layout(guides = "collect")
```


## Results

\color{violet} \textbf{Mean Width:}
```{r}
#| label: fig-width
#| echo: false
#| message: false

width_data_smimodel_ppr <- readRDS(here::here("results/SMIModel_PPR/plotData_combined_width_heat_smimodel_ppr.rds"))
width_data_backward <- readRDS(here::here("results/Backward/plotData_combined_width_heat_backward.rds"))
width_data_gaim <- readRDS(here::here("results/GAIM/plotData_combined_width_heat_gaim.rds"))
width_data_ppr <- readRDS(here::here("results/PPR/plotData_combined_width_heat_ppr.rds"))

p5 <- width_data_smimodel_ppr |>
  mutate(method = factor(method, levels = methods)) |>
  ggplot(aes(x = horizon, y = width, group = method, colour = method)) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  scale_colour_manual(values = cols) +
  xlab("Forecast horizon") +
  ylab("Mean width") +
  ylim(c(30, 80)) +
  ggtitle("SMI Model - PPR") +
  theme_bw() +
  theme(axis.title.x = element_blank()) +
  scale_x_discrete(breaks = seq(1, 92, by = 5))

p6 <- width_data_backward |>
  mutate(method = factor(method, levels = methods)) |>
  ggplot(aes(x = horizon, y = width, group = method, colour = method)) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  scale_colour_manual(values = cols) +
  xlab("Forecast horizon") +
  ylab("Mean width") +
  ylim(c(30, 80)) +
  ggtitle("Backward") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  scale_x_discrete(breaks = seq(1, 92, by = 5))

p7 <- width_data_gaim |>
  mutate(method = factor(method, levels = methods)) |>
  ggplot(aes(x = horizon, y = width, group = method, colour = method)) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  scale_colour_manual(values = cols) +
  xlab("Forecast horizon") +
  ylab("Mean width") +
  ylim(c(30, 80)) +
  ggtitle("GAIM") +
  theme_bw() +
  scale_x_discrete(breaks = seq(1, 92, by = 5))

p8 <- width_data_ppr |>
  mutate(method = factor(method, levels = methods)) |>
  ggplot(aes(x = horizon, y = width, group = method, colour = method)) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  scale_colour_manual(values = cols) +
  xlab("Forecast horizon") +
  ylab("Mean width") +
  ylim(c(30, 80)) +
  ggtitle("PPR") +
  theme_bw() +
  theme(axis.title.y = element_blank()) +
  scale_x_discrete(breaks = seq(1, 92, by = 5))

patchwork2 <- (p5 + p6) / (p7 + p8) & theme(legend.position = "bottom", 
                                           #legend.box = "vertical",
                                           legend.title = element_blank(),
                                           axis.text.x = element_text(size = 7),
                                           axis.title = element_text(size = 8),
                                           title = element_text(size = 8))
patchwork2 +
  plot_layout(guides = "collect")
```


## Summary

::: {.callout-note}
## \color{violet} Summary of Results (work-in-progress):

\begin{itemize}
    \item \textbf{Block Bootstrap} -- Under-coverage; too narrow
    \item \textbf{Conformal Prediction} -- Better achieves a target coverage, with acceptable sharpness 
  \end{itemize}
:::

::: {.callout-warning}
## \color{violet} Limitations:

\begin{itemize}
    \item \small Test set is not long enough for larger forecast horizons
    \item \small Hyper-parameter choices
  \end{itemize}
:::


# Future Work

## Future Work

\begin{block}{}
  \begin{itemize}
    \item We propose a novel methodology: \alert{Conformal Block Bootstrap (CBB)} \newline
    \begin{itemize}
      \item \textbf{\color{blue}A natural integration of BB and SCP} \newline
      \item \textbf{\color{blue}Exploits the strengths of both the methods} \newline
    \end{itemize}
  \end{itemize}
\end{block}

\only<2->{\begin{textblock}{10}(2.9, 5)
\fontsize{12}{16}\sf
\begin{block}{Find me :}
\faHome \hspace*{0.2cm} \href{https://nuwanipalihawadana.netlify.app/}{nuwanipalihawadana.netlify.app} \newline
\faLinkedin \hspace*{0.3cm} \href{https://www.linkedin.com/in/nuwani-palihawadana/}{in/nuwani-palihawadana} \newline
\faGithub \hspace*{0.3cm} \href{https://github.com/nuwani-palihawadana}{@nuwani-palihawadana} \newline
\faEnvelope \hspace*{0.3cm} \href{mailto:nuwani.kodikarapalihawadana@monash.edu}{nuwani.kodikarapalihawadana@monash.edu}
\end{block}
\end{textblock}}


## References

\begin{itemize}
\item Vovk, V., Gammerman, A., and Shafer, G. (2005), \textit{Algorithmic learning in a random world}, New York, NY: Springer.
\item Tibshirani, R., Barber, R., Candès, E., and Ramdas, A. (2019), “Conformal prediction under covariate shift”, \textit{Advances in neural information processing systems}, 2526--2536.
\item Barber, R. F., Candès, E. J., Ramdas, A., and Tibshirani, R. J. (2023), “Conformal prediction beyond exchangeability”, \textit{The Annals of Statistics}, 51, 816--845.
\item Gibbs, I., and Candès, E. (2021), “Adaptive conformal inference under distribution shift”, \textit{Advances in neural information processing systems}, 1660--1672.
\item Wang, X., and Hyndman, R. J. (2024), “Online conformal inference for multi-step time series forecasting”.
\end{itemize}
